Working with the Poisson distribution

### 2021-09-08 text/21/poisson-inferences.Rmd
More work on poisson-inferences.md

```{r setup}
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(magrittr)))
```

@@

@@

@@

Divide by T to get a confidence interval for $\lambda$.

The mathematical approach used here is known as the delta method and is [documented in general terms] on Wikipedia.

@@

```{r}
x <- 17
n <- 1989
rate <- x/n
rate
```

@@

To construct a confidence interval for the rate, first start with a confidence interval for the count rather than the rate.

```{r}
lo <- 
hi <- x+1.96*sqrt(x)
count_ci <- c(x-1.96*sqrt(x), x+1.96*sqrt(x))
count_ci
```

@@

by first getting a confidence interval for

$\frac{\lambda_1 T_1}{\lambda_2 T_2}$

and then dividing the interval by

$\frac{T_1}{T_2}$

@@

$ln\big(\frac{X_1}{X_2}\big)=ln(X_1)-ln(X_2)$

@@

$ln(X_1)-ln(X_2) \pm z_{\alpha/2}\sqrt{\frac{1}{X_1}+\frac{1}{X_2}}$

@@

$\frac{\lambda_1 T_1}{\lambda_2 T_2}$

Then exponentiate both sides of the interval and adjust by divinding by $T_1/T_2$.

@@

In a [2018 article][kuma1] published in the Indian Journal of Dermatology, the researchers compared the incidence rates of leprosy in different age groups. In patients under the age of 14 there were 33 cases cumulated across 290508.9 patient years. In patients 15-29, there were 52 cases 141,112.0 patient years. Here are the calculations for the confidence interval for the rate ratio.

First calculate the rates and their ratio. Since rates are so low, adjust them to a rate per 10,000 patient years.

```{r}
x1 <- 33
T1 <- 290508.9
x2 <- 52
T2 <- 141112
d<- 10000
x2 <- 60
T2 <- 100212.3

x1 <- 2
T1 <- 0.6*2
x2 <- 8
T2 <- 5.47*8
d <- 1

x1 <- 30
T1 <- 54308.7
x2 <- 60
T2 <- 51477.5
d <- 100000

x1 <- 215
T1 <- 4770
x2 <- 84
T2 <- 4193
d <- 1000

@@

Calculate the interval for $ln(\lambda_1 T_1/\lambda_2 T_2)$.

@@

Adjust for the patient years. Don't forget to round your final answer.

```{r}
ci3 <- ci2/(T1/T2)
round(ci3, 2)
```

@@

>>>>

```{r setup}
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(magrittr)))
```

@@

The mathematical approach used here is known as the delta method and is [documented in general terms][delt1] on Wikipedia with more specifics [here][delt2].

@@

Because logarithms behave the way they do, we know that

$ln(\lambda T)=ln(\lambda)-ln(T)$

so you convert the above interval to a confidence interval for $ln(\lambda)$ by subtracting $ln(T)$. This produces the interval

$ln(R) \pm z_{\alpha/2} \sqrt{\frac{1}{X}}$

@@

The interesting thing about logarithms is that if you change the units of time (hundreds of patient years to thousands of patient years, for example), the width of the interval on the log scale is still the same.

@@

```{r}
x <- 17
n <- 1989

x <- 299
n <- 8963

x <- 84
n <- 4193

x <- 215
n <- 4770

rate <- x/n
rate
```

@@

To construct a confidence interval for the rate, first start with a confidence interval using the log rate.

```{r}
count_ci <- c(x-1.96*sqrt(x), x+1.96*sqrt(x))
count_ci
```

@@

by using the same sort of calculations.

@@

$ln\big(\frac{R_1}{R_2}\big)=ln(R_1)-ln(R_2)$

@@

$ln(R_1)-ln(R_2) \pm z_{\alpha/2}\sqrt{\frac{1}{X_1}+\frac{1}{X_2}}$

@@

$ln(\frac{\lambda_1}{\lambda_2})$

Then exponentiate both sides of the interval.

@@

In a [2020 article][hess1] published in the Journal of Athletic Training, the researchers compared the incidence rates of injuries for professional ultimate frisbee players during games and during practice.

First calculate the rates and their ratio. Since rates are so low, adjust them to a rate per 1,000 patient years.

```{r}
x1 <- 215
T1 <- 4770
x2 <- 84
T2 <- 4193
d <- 1000

@@

Calculate the interval on the log scale.

@@

Don't forget to round your final answer.

```{r}
round(ci2, 2)
```

This interval does not match the one published in the paper because the researchers adjusted for factors like field type.

```{r}
injuries <- data.frame(x=c(x1, x2), T=c(T1, T2), g=1:0)
glm(x~g, family=poisson, offset=log(T), data=injuries) %>%
  tidy -> c
c
```

@@

[hess1]: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7017898/
